Efficient ancestry and mutation simulation with msprime 1.0

### 2021-08-11 paper.tex
Adding AlphaSimR link/citation

Part of the reason for this poor record of software reuse and modularity is the
lack of standardised file formats, and in particular, the absence of common
library infrastructure to abstract the details of interchanging simulation
data. Although \msprime\ also supports simulating both ancestry and mutations,
the two aspects are functionally independent within the software.
Both ancestry and mutation simulators are present in \msprime\
for reasons of convenience and history,
and could be split into separate packages.
The mature and efficient C and Python interfaces for
\tskit\ make it straightforward to add further information to
an existing file, and because of its efficient data interchange mechanisms,
there is no performance penalty for performing additional operations
in a different Python package.
Thanks to this interoperability, \msprime's
mutation generator can work with \emph{any} \tskit\ tree sequence,
be it simulated using \SLiM~\citep{haller2019slim} or
\fwdpy~\citep{thornton2014cpp}, or estimated from real
data~\citep{kelleher2019inferring,speidel2019method,wohns2021unified}.
It is a modular
component intended to fit into a larger software ecosystem, and
is in no way dependent on \msprime's ancestry simulator.

@@

\subsection*{Integration with forward simulators}
A unique feature of \msprime\ is its ability to simulate
genetic ancestries by extending an existing partial
genetic ancestry. Given a tree sequence that
is complete up until time $t$ as input
(where marginal trees may or may not have fully coalesced),
\msprime\ can efficiently obtain the segments of ancestral material
present at this time, and then run the simulation backwards in time from there.
This allows a simulated ancestry to be produced by any
number of different processes across disjoint time slices.
In practice this feature is used to ``complete''
forwards-time ancestry simulations~\citep{kelleher2018efficient}
that may have not fully coalesced. This process
(``recapitation'') can be orders of magnitude faster than
the standard approach of neutral burn-in; see
\cite{haller2018tree} for more details and examples.
This interoperability between simulators, where a partial ancestry
simulation produced by \SLiM~\citep{haller2019slim}
or \fwdpy~\citep{thornton2014cpp} can be picked up and completed
by another simulator, with complete information
retained---at scale---is unprecedented.

>>>>

Part of the reason for this poor record of software reuse and modularity is the
lack of standardised file formats, and in particular, the absence of common
library infrastructure to abstract the details of interchanging simulation
data. Although \msprime\ also supports simulating both ancestry and mutations,
the two aspects are functionally independent within the software.
Both ancestry and mutation simulators are present in \msprime\
for reasons of convenience and history,
and could be split into separate packages.
The mature and efficient C and Python interfaces for
\tskit\ make it straightforward to add further information to
an existing file, and because of its efficient data interchange mechanisms,
there is no performance penalty for additional operations
in a different Python package.
Thanks to this interoperability, \msprime's
mutation generator can work with \emph{any} \tskit\ tree sequence,
be it simulated using \SLiM~\citep{haller2019slim} or
\fwdpy~\citep{thornton2014cpp}, or estimated from real
data~\citep{kelleher2019inferring,speidel2019method,wohns2021unified}.
It is a modular
component intended to fit into a larger software ecosystem, and
is in no way dependent on \msprime's ancestry simulator.

@@

\subsection*{Integration with forward simulators}
A unique feature of \msprime\ is its ability to simulate
genetic ancestries by extending an existing partial
genetic ancestry. Given a tree sequence that
is complete up until time $t$ as input
(where marginal trees may or may not have fully coalesced),
\msprime\ can efficiently obtain the segments of ancestral material
present at this time, and then run the simulation backwards in time from there.
This allows a simulated ancestry to be produced by any
number of different processes across disjoint time slices.
In practice this feature is used to ``complete''
forwards-time ancestry simulations~\citep{kelleher2018efficient}
that may have not fully coalesced. This process
(``recapitation'') can be orders of magnitude faster than
the standard approach of neutral burn-in; see
\cite{haller2018tree} for more details and examples.
This interoperability between simulators, where a partial ancestry
simulation produced by \SLiM~\citep{haller2019slim}
or \fwdpy~\citep{thornton2014cpp} can be picked up and completed
by another simulator, with complete information
retained---at scale---is unprecedented.
This interoperability is indicating an opportunity for
other forward genetic simulators ~\citep[e.g.][]{Gaynor2021AlphaSimR}
to leverage tree sequence data format and associated tools.
